### 2020-04-04 thesis/content/chapter-ltd.tex
various fixes + started with 2-wl-gnn implementation section

The $k$-\acs{gnn} is a \ac{gcnn} inspired by the $k$-\ac{wl} algorithm, it convolves feature vectors of vertex $k$-multisets.
In this section we will compare its $k = 2$ variant with the 2-\ac{wl} algorithm.
The main difference between the two boils down to their notion of ``neighborhood''.
As already briefly mentioned in \cref{sec:related:gcr:nn}, 2-\acsp{gnn} define the neighbors of an edge $\textcolor{t_blue}{e_{ij}} = (v_i, v_j)$ to be the edges that are incident to either $v_i$ or $v_j$. % chktex 25
In 2-\acs{wl} on the other hand, the neighbors of $\textcolor{t_blue}{e_{ij}}$ are the edge pairs ${\left\{(\textcolor{t_red}{e_{il}}, \textcolor{t_darkgreen}{e_{lj}})\right\}}_{v_l \in \mathcal{V}_G}$, i.e.\ all possible paths of length two that start at $v_i$ and end at $v_j$. % chktex 21 chktex 25
This difference becomes clear when comparing the definition of convolution in 2-\acsp{gnn} with that of color refinement in 2-\acs{wl} (see \cfullref[ on ]{eq:related:kgnn-layer} and \cfullref[ on ]{defn:related:wlk-refine}):
\begin{alignat*}{4}
	\text{2-\acs{gnn}\footnotemark: } && Z_G^{(t)}[e_{ij}] &= \sigma\Biggl( \textcolor{t_blue}{Z_G^{(t-1)}[e_{ij}]} W^{(t)} + &&\left( \smashoperator[r]{\sum_{v_l \in \Gamma_G(v_j)}} \textcolor{t_red}{Z_G^{(t-1)}[e_{il}]} + \smashoperator[lr]{\sum_{v_l \in \Gamma_G(v_i)}} \textcolor{t_darkgreen}{Z_G^{(t-1)}[e_{lj}]} \right) W_{\Gamma}^{(t)} \Biggr) \\ % chktex 25
	\text{2-\acs{wl}: } && \chi_{G,2}^{(t)}(e_{ij}) &= h\Bigl(\textcolor{t_blue}{\chi_{G,2}^{(t-1)}(e_{ij})}, &&\ldblbrace (\textcolor{t_red}{\chi_{G,2}^{(t-1)}(e_{il})}, \textcolor{t_darkgreen}{\chi_{G,2}^{(t-1)}(e_{lj})})\, |\, v_l \in \mathcal{V}_G \rdblbrace\Bigr) % chktex 25
\end{alignat*}\footnotetext{
	To highlight the relation between 2-\acsp{gnn} and 2-\acs{wl}, a 2-\acs{gnn} definition that is not generally correct is shown here;
	for self-loops with $i = j$ it incorrectly sums the feature vectors of neighboring edges twice.
	The correct general formula uses a single sum over $v_l \in {\Gamma_G(v_j) \cup \Gamma_G(v_i)}$.
}\begin{figure}[ht]
	\centering
	\includegraphics[width=\linewidth]{gfx/ltd/2gnn-2wl-diff.pdf}
	\caption[Edge colorings on which 2-\acsp{gnn} and 2-\acs{wl} behave differently.]{
		Two edge colorings on which 2-\acsp{gnn} and 2-\acs{wl} behave differently.
		A 2-\acs{gnn} will refine the ``color vector'' of $e_{ij}$ to \textsf{\textbf{\small\textcolor{t_blue}{D}}} for both initial colorings.
		2-\acs{wl} on the other hand differentiates both colorings by preserving the color tuple information.
	}\label{fig:ltd:2gnn-2wl-diff}
\end{figure}

@@

	W.l.o.g.\ we define the initial colors of the loop vertices as $\chi^{(0)}(v) = \textcolor{t_blue}{\texttt{A}}$ for all $v \in L_G \cup L_H$. % chktex 25
	The initial colors of the edge vertices are defined as $\chi^{(0)}(e) = \textcolor{t_red}{\texttt{B}}$ for all $e \in E_G \cup E_H$. % chktex 25
	Note that each loop vertex $\ldblbrace v_i, v_i \rdblbrace$ with $v_i \in \mathcal{V}_G \cup \mathcal{V}_H$ has $d$ neighbors, the edges incident to $v_i$.
	Similarly each edge vertex $\ldblbrace v_i, v_j \rdblbrace$ has $2d$ neighbors, two of which are the loop vertices $\ldblbrace v_i, v_i \rdblbrace$ and $\ldblbrace v_j, v_j \rdblbrace$ with the remaining $2d - 2$ neighbors corresponding to the edges that are incident to $e_{ij}$.
	This is illustrated in \cref{fig:ltd:edge-neighborhood}.

	After one color refinement step we get $\chi^{(1)}(v) = h(\textcolor{t_blue}{\texttt{A}}, \ldblbrace \underbrace{\textcolor{t_red}{\texttt{B}}, \dots, \textcolor{t_red}{\texttt{B}}}_{d\text{ times}} \rdblbrace) \eqqcolon \textcolor{t_blue}{\texttt{C}}$ for all loop vertices $v \in  L_G \cup L_H$ and $\chi^{(1)}(e) = h(\textcolor{t_red}{\texttt{B}}, \ldblbrace \textcolor{t_blue}{\texttt{A}}, \textcolor{t_blue}{\texttt{A}}, \underbrace{\textcolor{t_red}{\texttt{B}}, \dots, \textcolor{t_red}{\texttt{B}}}_{\mathclap{2d - 2 \text{ times}}} \rdblbrace) \eqqcolon \textcolor{t_red}{\texttt{D}}$. % chktex 25
	This means that $\chi^{(0)}$ and $\chi^{(1)}$ are identical up to the color substitutions $\textcolor{t_blue}{\texttt{A}} \to \textcolor{t_blue}{\texttt{C}}$ and $\textcolor{t_red}{\texttt{B}} \to \textcolor{t_red}{\texttt{D}}$, i.e.\ $\chi^{(0)} \equiv \chi^{(1)}$, which in turn implies that 1-\ac{wl} terminates after one iteration. % chktex 25
	\Cref{lem:ltd:wl1-regular-edge-neighbor-limit} then directly follows since both $G^{\mathcal{E}}$ and $H^{\mathcal{E}}$ have $n$ vertices with the final color $\textcolor{t_blue}{\texttt{C}}$ and $\frac{nd}{2}$ vertices with the final color $\textcolor{t_red}{\texttt{D}}$, i.e.\ $G^{\mathcal{E}} \mathrel{\simeq_1} H^{\mathcal{E}}$. % chktex 25
\end{proof}
\begin{figure}[ht]
	\centering
	\includegraphics[width=\linewidth]{gfx/ltd/edge-neighborhood.pdf}
	\caption{
		Illustration of the edge neighborhood graphs of two 2-regular graphs of size 6.
	}\label{fig:ltd:edge-neighborhood}
\end{figure}
\begin{prop}\label{prop:ltd:2gnn-regular-limit}
	A 2-\acs{gnn} cannot distinguish regular graphs of the same size and therefore has a lower discriminative power than 2-\ac{wl}.
\end{prop}
\begin{proof}
	The proposition directly follows from \cref{prop:ltd:2gnn-wl1-limit}, \cref{lem:ltd:wl1-regular-edge-neighbor-limit} and the fact that 2-\acs{wl} is able to distinguish most regular graphs~\cite[cor.~1.8.6]{Immerman1990}.
\end{proof}

@@

Based on the 2-multiset and the neighborhood localization simplifications we can now define the 2-\ac{wl} convolution operator.

@@

@@

In this section we will analyze the discriminative and computational power of the 2-\ac{wl} convolution operator that was just defined.
We begin by showing that 2-\acs{wl}-\acsp{gnn} are at least as powerful as 1-\acs{wl}.

@@

@@

@@

This concludes our analysis of the expressive power of 2-\acs{wl}-\acsp{gnn}.
The key takeaway from this section is that 2-\acs{wl}-\acsp{gnn} are more powerful than all vertex neighborhood aggregation \acp{gcnn} because the power of the latter is upper bounded by 1-\acs{wl}.
Additionally we can conclude that 2-\acs{wl}-\acsp{gnn} are in fact also more powerful than 2-\acsp{gnn} due to \cref{prop:ltd:2gnn-regular-limit}.
A more thorough discussion of the relation between 2-\acs{wl}-\acsp{gnn} and 2-\acs{wl} is left for future work.

@@

>>>>

The $k$-\acs{gnn} is a \ac{gcnn} inspired by the $k$-\ac{wl} algorithm, it convolves feature vectors of vertex $k$-multisets.
In this section we will compare its $k = 2$ variant with the 2-\ac{wl} algorithm.
The main difference between the two boils down to their notion of ``neighborhood''.
As already briefly mentioned in \cref{sec:related:gcr:nn}, 2-\acsp{gnn} define the neighbors of an edge $\textcolor{t_blue}{e_{ij}} = (v_i, v_j)$ to be the edges that are incident to either $v_i$ or $v_j$. % chktex 25
In 2-\acs{wl} on the other hand, the neighbors of $\textcolor{t_blue}{e_{ij}}$ are the edge pairs ${\left\{(\textcolor{t_red}{e_{il}}, \textcolor{t_darkgreen}{e_{lj}})\right\}}_{v_l \in \mathcal{V}_G}$, i.e.\ all possible paths of length two that start at $v_i$ and end at $v_j$. % chktex 21 chktex 25
This difference becomes clear when comparing the definition of convolution in 2-\acsp{gnn} with that of color refinement in 2-\acs{wl} (see \cfullref[ on ]{eq:related:kgnn-layer} and \cfullref[ on ]{defn:related:wlk-refine}):
\begin{alignat*}{4}
	\text{2-\acs{gnn}\footnotemark: } && Z_G^{(t)}[e_{ij}] &= \sigma\Biggl( \textcolor{t_blue}{Z_G^{(t-1)}[e_{ij}]} W^{(t)} + &&\left( \smashoperator[r]{\sum_{v_l \in \Gamma_G(v_j)}} \textcolor{t_red}{Z_G^{(t-1)}[e_{il}]} + \smashoperator[lr]{\sum_{v_l \in \Gamma_G(v_i)}} \textcolor{t_darkgreen}{Z_G^{(t-1)}[e_{lj}]} \right) W_{\Gamma}^{(t)} \Biggr) \\ % chktex 25
	\text{2-\acs{wl}: } && \chi_{G,2}^{(t)}(e_{ij}) &= h\Bigl(\textcolor{t_blue}{\chi_{G,2}^{(t-1)}(e_{ij})}, &&\ldblbrace (\textcolor{t_red}{\chi_{G,2}^{(t-1)}(e_{il})}, \textcolor{t_darkgreen}{\chi_{G,2}^{(t-1)}(e_{lj})})\, |\, v_l \in \mathcal{V}_G \rdblbrace\Bigr) % chktex 25
\end{alignat*}\footnotetext{
	To highlight the relation between 2-\acsp{gnn} and 2-\acs{wl}, a 2-\acs{gnn} definition that is not generally correct is shown here;
	for self-loops with $i = j$ it incorrectly sums the feature vectors of neighboring edges twice.
	The correct general formula uses a single sum over $v_l \in {\Gamma_G(v_j) \cup \Gamma_G(v_i)}$.
}\begin{figure}[ht]
	\centering
	\includegraphics[width=\linewidth]{gfx/ltd/2gnn-2wl-diff.pdf}
	\caption[Edge colorings on which 2-\acsp{gnn} and 2-\acs{wl} behave differently.]{
		Two edge colorings on which 2-\acsp{gnn} and 2-\acs{wl} behave differently.
		A 2-\acs{gnn} will refine the ``color vector'' of $e_{ij}$ to \colorlabel{t_blue}{D} for both initial colorings.
		2-\acs{wl} on the other hand differentiates both colorings by preserving the color tuple information.
	}\label{fig:ltd:2gnn-2wl-diff}
\end{figure}

@@

	W.l.o.g.\ we define the initial colors of the loop vertices as $\chi^{(0)}(v) = \colorlabel{t_blue}{A}$ for all $v \in L_G \cup L_H$. % chktex 25
	The initial colors of the edge vertices are defined as $\chi^{(0)}(e) = \textcolor{t_red}{\texttt{B}}$ for all $e \in E_G \cup E_H$. % chktex 25
	Note that each loop vertex $\ldblbrace v_i, v_i \rdblbrace$ with $v_i \in \mathcal{V}_G \cup \mathcal{V}_H$ has $d$ neighbors, the edges incident to $v_i$.
	Similarly each edge vertex $\ldblbrace v_i, v_j \rdblbrace$ has $2d$ neighbors, two of which are the loop vertices $\ldblbrace v_i, v_i \rdblbrace$ and $\ldblbrace v_j, v_j \rdblbrace$ with the remaining $2d - 2$ neighbors corresponding to the edges that are incident to $e_{ij}$.
	This is illustrated in \cref{fig:ltd:edge-neighborhood}.

	After one color refinement step we get $\chi^{(1)}(v) = h(\colorlabel{t_blue}{A}, \ldblbrace \underbrace{\textcolor{t_red}{\texttt{B}}, \dots, \textcolor{t_red}{\texttt{B}}}_{d\text{ times}} \rdblbrace) \eqqcolon \textcolor{t_blue}{\texttt{C}}$ for all loop vertices $v \in  L_G \cup L_H$ and $\chi^{(1)}(e) = h(\textcolor{t_red}{\texttt{B}}, \ldblbrace \colorlabel{t_blue}{A}, \colorlabel{t_blue}{A}, \underbrace{\textcolor{t_red}{\texttt{B}}, \dots, \textcolor{t_red}{\texttt{B}}}_{\mathclap{2d - 2 \text{ times}}} \rdblbrace) \eqqcolon \textcolor{t_red}{\texttt{D}}$. % chktex 25
	This means that $\chi^{(0)}$ and $\chi^{(1)}$ are identical up to the color substitutions $\colorlabel{t_blue}{A} \to \textcolor{t_blue}{\texttt{C}}$ and $\textcolor{t_red}{\texttt{B}} \to \textcolor{t_red}{\texttt{D}}$, i.e.\ $\chi^{(0)} \equiv \chi^{(1)}$, which in turn implies that 1-\ac{wl} terminates after one iteration. % chktex 25
	\Cref{lem:ltd:wl1-regular-edge-neighbor-limit} then directly follows since both $G^{\mathcal{E}}$ and $H^{\mathcal{E}}$ have $n$ vertices with the final color $\textcolor{t_blue}{\texttt{C}}$ and $\frac{nd}{2}$ vertices with the final color $\textcolor{t_red}{\texttt{D}}$, i.e.\ $G^{\mathcal{E}} \mathrel{\simeq_1} H^{\mathcal{E}}$. % chktex 25
\end{proof}
\begin{figure}[ht]
	\centering
	\includegraphics[width=\linewidth]{gfx/ltd/edge-neighborhood.pdf}
	\caption{
		Illustration of the edge neighborhood graphs of two 2-regular graphs of size 6.
	}\label{fig:ltd:edge-neighborhood}
\end{figure}
\begin{prop}\label{prop:ltd:2gnn-regular-limit}
	A 2-\acs{gnn} cannot distinguish regular graphs of the same size and therefore has a lower discriminative power than 2-\ac{wl}.
\end{prop}
\begin{proof}
	The proposition directly follows from \cref{prop:ltd:2gnn-wl1-limit}, \cref{lem:ltd:wl1-regular-edge-neighbor-limit} and the fact that 2-\acs{wl} is able to distinguish most regular graphs~\cite[cor.~1.8.6]{Immerman1990}.
\end{proof}

@@

Based on the 2-multiset and the neighborhood localization simplifications we can now define the 2-\ac{wl} convolution operator and the corresponding so-called \textit{2-\acs{wl}-\acs{gnn}}.

@@

		2-\acs{wl}-\acsp{gnn} do not generally have this problem because $\kappa^{(t)}(\colorlabel{t_blue}{A}, \ldblbrace \colorlabel{t_red}{B}, \colorlabel{t_red}{B} \rdblbrace) + \kappa^{(t)}(\colorlabel{t_blue}{A}, \ldblbrace \colorlabel{t_darkgreen}{C}, \colorlabel{t_darkgreen}{C} \rdblbrace) \neq \kappa^{(t)}(\colorlabel{t_blue}{A}, \ldblbrace \colorlabel{t_red}{B}, \colorlabel{t_darkgreen}{C} \rdblbrace) + \kappa^{(t)}(\colorlabel{t_blue}{A}, \ldblbrace \colorlabel{t_red}{B}, \colorlabel{t_darkgreen}{C} \rdblbrace)$ if $\sigma_{\Gamma}$ is chosen to be a nonlinear activation function. % chktex 25

@@

In this section we will analyze the power of \acp{gnn} which use the 2-\acs{wl} convolution operator that was just defined.
Our goal is to show that such 2-\acs{wl}-\acsp{gnn} have a strictly larger discriminative power than 1-\acs{wl}.
We begin by proving that 2-\acs{wl}-\acsp{gnn} are at least as powerful as 1-\acs{wl}.

@@

	\begin{gather*}\allowdisplaybreaks

@@

		\end{aligned}\\[10pt]

@@

This concludes our analysis of the expressive power of 2-\acs{wl}-\acsp{gnn}.
The key takeaway from this section is that 2-\acs{wl}-\acsp{gnn} are more powerful than all vertex neighborhood aggregation \acp{gcnn} because the power of the latter is upper bounded by 1-\acs{wl}.
Additionally we can conclude that 2-\acs{wl}-\acsp{gnn} are in fact also more powerful than 2-\acsp{gnn} due to \cref{prop:ltd:2gnn-regular-limit}.

Note that no statement regarding the discriminative or computational power of 2-\acs{wl}-\acsp{gnn} compared to 2-\acs{wl} was made.
It is easy to see that 2-\acs{wl}-\acsp{gnn} \textit{generally} cannot have the same power as 2-\acs{wl} due to the neighborhood localization simplification;
e.g.\ for a small neighborhood radius of $r = 1$, nonexistent edges $e_{ij} \notin \mathcal{E}_G$ do not have a feature vector which implies that the proof of 2-\acs{wl}'s $m$-cycle counting ability no longer holds for $m \geq 4$ (see \cfullref{prop:related:wl2-cycle-count}).
We leave a more thorough discussion of the relation between 2-\acs{wl}-\acsp{gnn} and 2-\acs{wl} for future work.

@@

Apart from the theoretical expressive power of a model, it also has to be computable efficiently in order to be useful in practice.
In this section we will therefore describe how 2-\acs{wl} convolutions can be implemented on \acp{gpgpu}.

Efficient high-level software libraries for the implementation of vertex neighborhood convolution approaches such as \ac{gcn} or \ac{gin} already exist.
They describe convolutions via a message-passing abstraction in which vertex feature vectors are passed along their neighboring edges (see \citet{Battaglia2018}).
A few notable implementations of this abstraction are the \citetitle{GN} library~\cite{GN}, \citetitle{PyG}~\cite{Fey2019}\cite{PyG}, \citetitle{DGL}~\cite{Wang2019}\cite{DGL} and \citetitle{Spektral}~\cite{Spektral}.

### 2020-04-04 thesis/content/chapter-related-work.tex
various fixes + started with 2-wl-gnn implementation section

In order to process a graph $G$ of size $n = |\mathcal{V}_G|$, one is forced to choose some encoding, e.g.\ an adjacency matrix $A_G \in {\{0, 1\}}^{n \times n}$.
Such an encoding introduces a vertex ordering $v_1, \dots, v_n$ that does not carry any semantic meaning.
Consequently there are $n!$ equivalent encodings of $G$. % chktex 40
To represent those encodings we introduce the notion of ordered induced subgraphs.
\begin{defn}\label{defn:related:ordered-subgraph}
	For all vertex $k$-tuples $v = (v_1, \dots, v_k) \in \mathcal{V}_G^k$ let $\hat{v} = \set(v) \coloneqq \{ v_i\, |\, i \in [k] \}$.
	Then $G[v] \coloneqq (\hat{v}, \mathcal{E}_G \cap \hat{v}^2, v)$ is called the \textit{ordered subgraph} of $G$ induced by $v$.
	We denote the set of all ordered subgraphs with $\mathcal{G}_{\mathit{ord}}$.
	If $\hat{v} = \mathcal{V}_G$ and $k = |\mathcal{V}_G|$, we call $v$ an \textit{ordering} of $G$ and $G[v]$ an \textit{encoding} of $G$.
\end{defn}
\begin{defn}\label{defn:related:ordered-subgraph-equivalence}
	Two ordered induced subgraphs $G[v]$ and $H[w]$ with $(v_1, \dots, v_k) \in \mathcal{V}_G^k$ and $(w_1, \dots, w_k) \in \mathcal{V}_H^k$ are \textit{equivalent} ($G[v] \equiv H[w]$) iff.\
	\begin{align*}
		&\quad\forall i, j \in [k]: (v_i, v_j) \in \mathcal{E}_G \leftrightarrow (w_i, w_j) \in \mathcal{E}_H \\
		&\, \begin{aligned}
			\land&\ \forall i, j \in [k]: l_G[v_i, v_j] = l_H[w_i, w_j] &\land&\quad\forall i \in [k]: l_G[v_i] = l_H[w_i] \\
			\land&\ \forall i, j \in [k]: x_G[v_i, v_j] = x_H[w_i, w_j] &\land&\quad\forall i \in [k]: x_G[v_i] = x_H[w_i]\text{.}
		\end{aligned}
	\end{align*}
	Using this notion of equivalence, we call ${\left[ G[v] \right]} \coloneqq \{ G[w]\, |\, G[w] \equiv G[v] \land w \in \mathcal{V}_G^{*} \}$ the \textit{ordered subgraph equivalence class} of $G[v]$.
\end{defn}
\begin{defn}
	Two graphs $G$ and $H$ are \textit{isomorphic} ($G \simeq H$) iff.\ there are equivalent encodings $G[v] \equiv H[w]$ of them.
	Consequently $[G] \coloneqq \{ H\, |\, H \simeq G \}$ is called the \textit{isomorphism class} of $G$.
\end{defn}
The goal of the \ac{gi} problem is to check whether $G \simeq H$ for two arbitrary graphs.
Even though there is no known universal polynomial algorithm that solves \ac{gi}, \citet{Babai1980} showed that almost all graphs can be trivially distinguished in linear time.
More recently \citet{Babai2015} also presented a quasipolynomial upper time bound for the remaining hard \ac{gi} instances.
For all practical purposes, \ac{gi} can therefore be solved efficiently; e.g.\ via the \texttt{nauty} program~\cite{McKay}\cite{McKay2013}.
One important subroutine in most \ac{gi} checkers is the \acl{wl} algorithm which will be described next.

@@

\subsubsection{The 1-dimensional \acs{wl} algorithm}
In the 1-dimensional \ac{wl} algorithm (1-WL), a color is assigned to each vertex of a graph.
If the vertices $v \in \mathcal{V}_G$ of the input graph $G$ are labeled, those labels $l_G[v] \in L_{\mathcal{V}} \subseteq \mathcal{C}$ can be used as the initial graph coloring $\chi_{G,1}^{(0)}(v) \coloneqq l_G[v]$.
Since \ac{wl} colors are inherently discrete, continuous vertex feature vectors $x_G[v]$ are not considered here.
For unlabeled graphs a constant coloring is used, e.g.\ $\forall v \in \mathcal{V}_G: \chi_{G,1}^{(0)}(v) = \textcolor{t_blue}{\texttt{A}}$ for some initial color $\textcolor{t_blue}{\texttt{A}} \in \mathcal{C}$. % chktex 25
In each iteration of the 1-\acs{wl} color refinement algorithm, the following neighborhood aggregation scheme is used to compute a new color for all vertices:
\begin{defn}\label{defn:related:wl1-refine}
	$\chi_{G,1}^{(i+1)}(v) \coloneqq h\left(\chi_{G,1}^{(i)}(v), \ldblbrace \chi_{G,1}^{(i)}(u)\, |\, u \in \Gamma_{G}(v) \rdblbrace\right)$,
	with $\Gamma_G(v)$ denoting the set of adjacent vertices of $v \in \mathcal{V}_G$ and $h: \mathcal{C}^* \to \mathcal{C}$ denoting an injective hash function that assigns a unique color to each finite combination of colors.
\end{defn}
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.72\linewidth]{gfx/related-work/wl1-refine.pdf}
	\caption[Example 1-WL color refinement steps.]{
		Example 1-WL color refinement steps.
		After two iterations the coloring stabilizes.
		Each vertex $v$ is labeled with its current color and has its previous color and the colors of the hashed neighbors $\Gamma_G(v)$ written next to it (see \cref{defn:related:wl1-refine}).
	}\label{fig:related:wl1-refine}
\end{figure}
In practice the hash function $h$ is usually defined lazily by using $\mathcal{C} \subseteq \mathbb{N}$ and enumerating color combinations in whichever order they are hashed s.t.\ a new color is introduced every time a previously unseen color combination appears at runtime\footnote{
	$\mathcal{C}$ is countably infinite but we assume it only contains the finite no.\ of colors occurring in a dataset.
}.

@@

This simple approach usually requires a careful choice of subgraphs but can still be competitive with the other more recent approaches we will look at in the following sections.
Fingerprint embeddings are for example used in multiple state-of-the-art toxicity prediction tools like RASAR~\cite{Luechtefeld2018}\cite{ToxTrack}, the \ac{toxTest}~\cite{TEST} or ProTox~\cite{Drwal2014}\cite{ProTox}\cite{Banerjee2018}.

>>>>

In order to process a graph $G$ of size $n = |\mathcal{V}_G|$, one is forced to choose some encoding, e.g.\ an adjacency matrix $A_G \in {\{0, 1\}}^{n \times n}$.
Such an encoding introduces a vertex ordering $v_1, \dots, v_n$ that does not carry any semantic meaning.
Consequently there are $n!$ equivalent encodings of $G$. % chktex 40
To represent those encodings we introduce the notion of ordered induced subgraphs.
\begin{defn}\label{defn:related:ordered-subgraph}
	For all vertex $k$-tuples $v = (v_1, \dots, v_k) \in \mathcal{V}_G^k$ let $\hat{v} = \set(v) \coloneqq \{ v_i\, |\, i \in [k] \}$.
	Then $G[v] \coloneqq (\hat{v}, \mathcal{E}_G \cap \hat{v}^2, v)$ is called the \textit{ordered subgraph} of $G$ induced by $v$.
	We denote the set of all ordered subgraphs with $\mathcal{G}_{\mathit{ord}}$.
	If $\hat{v} = \mathcal{V}_G$ and $k = |\mathcal{V}_G|$, we call $v$ an \textit{ordering} of $G$ and $G[v]$ an \textit{encoding} of $G$.
\end{defn}
\begin{defn}\label{defn:related:ordered-subgraph-equivalence}
	Two ordered induced subgraphs $G[v]$ and $H[w]$ with $(v_1, \dots, v_k) \in \mathcal{V}_G^k$ and $(w_1, \dots, w_k) \in \mathcal{V}_H^k$ are \textit{equivalent} ($G[v] \equiv H[w]$) iff.\
	\begin{align*}
		&\quad\forall i, j \in [k]: (v_i, v_j) \in \mathcal{E}_G \leftrightarrow (w_i, w_j) \in \mathcal{E}_H \\
		&\, \begin{aligned}
			\land&\ \forall i, j \in [k]: l_G[v_i, v_j] = l_H[w_i, w_j] &\land&\quad\forall i \in [k]: l_G[v_i] = l_H[w_i] \\
			\land&\ \forall i, j \in [k]: x_G[v_i, v_j] = x_H[w_i, w_j] &\land&\quad\forall i \in [k]: x_G[v_i] = x_H[w_i]\text{.}
		\end{aligned}
	\end{align*}
	Using this notion of equivalence, we call ${\left[ G[v] \right]} \coloneqq \{ G[w]\, |\, G[w] \equiv G[v] \land w \in \mathcal{V}_G^{*} \}$ the \textit{ordered subgraph equivalence class} of $G[v]$.
\end{defn}
\begin{defn}
	Two graphs $G$ and $H$ are \textit{isomorphic} ($G \simeq H$) iff.\ there are equivalent encodings $G[v] \equiv H[w]$ of them.
	Consequently $[G] \coloneqq \{ H\, |\, H \simeq G \}$ is called the \textit{isomorphism class} of $G$.
\end{defn}
The goal of the \ac{gi} problem is to check whether $G \simeq H$ for two arbitrary graphs.
Even though there is no known universal polynomial algorithm that solves \ac{gi}, \citet{Babai1980} showed that almost all graphs can be trivially distinguished in linear time.
More recently \citet{Babai2015} also presented a quasipolynomial upper time bound for the remaining hard \ac{gi} instances.
For all practical purposes, \ac{gi} can therefore be solved efficiently; e.g.\ via the \texttt{nauty} program~\cite{McKay2013}\cite{McKay}.
One important subroutine in most \ac{gi} checkers is the \acl{wl} algorithm which will be described next.

@@

\subsubsection{The 1-dimensional \acs{wl} algorithm}
In the 1-dimensional \ac{wl} algorithm (1-WL), a color is assigned to each vertex of a graph.
If the vertices $v \in \mathcal{V}_G$ of the input graph $G$ are labeled, those labels $l_G[v] \in L_{\mathcal{V}} \subseteq \mathcal{C}$ can be used as the initial graph coloring $\chi_{G,1}^{(0)}(v) \coloneqq l_G[v]$.
Since \ac{wl} colors are inherently discrete, continuous vertex feature vectors $x_G[v]$ are not considered here.
For unlabeled graphs a constant coloring is used, e.g.\ $\forall v \in \mathcal{V}_G: \chi_{G,1}^{(0)}(v) = \colorlabel{t_blue}{A}$ for some initial color $\colorlabel{t_blue}{A} \in \mathcal{C}$. % chktex 25
In each iteration of the 1-\acs{wl} color refinement algorithm, the following neighborhood aggregation scheme is used to compute a new color for all vertices:
\begin{defn}\label{defn:related:wl1-refine}
	$\chi_{G,1}^{(i+1)}(v) \coloneqq h\left(\chi_{G,1}^{(i)}(v), \ldblbrace \chi_{G,1}^{(i)}(u)\, |\, u \in \Gamma_{G}(v) \rdblbrace\right)$,
	with $\Gamma_G(v)$ denoting the set of adjacent vertices of $v \in \mathcal{V}_G$ and $h: \mathcal{C}^* \to \mathcal{C}$ denoting an injective hash function that assigns a unique color to each finite combination of colors.
\end{defn}
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.72\linewidth]{gfx/related-work/wl1-refine.pdf}
	\caption[Example 1-WL color refinement steps.]{
		Example 1-WL color refinement steps.
		After two iterations the coloring stabilizes.
		Each vertex $v$ is labeled with its current color and has its previous color and the colors of the hashed neighbors $\Gamma_G(v)$ written next to it (see \cref{defn:related:wl1-refine}).
	}\label{fig:related:wl1-refine}
\end{figure}
In practice the hash function $h$ is usually defined lazily by using $\mathcal{C} \subseteq \mathbb{N}$ and enumerating color combinations in whichever order they are hashed s.t.\ a new color is introduced every time a previously unseen color combination appears at runtime\footnote{
	$\mathcal{C}$ is countably infinite but we assume it only contains the finite no.\ of colors occurring in a dataset.
}.

@@

This simple approach usually requires a careful choice of subgraphs but can still be competitive with the other more recent approaches we will look at in the following sections.
Fingerprint embeddings are for example used in multiple state-of-the-art toxicity prediction tools like RASAR~\cite{Luechtefeld2018}\cite{ToxTrack}, the \ac{toxTest}~\cite{TEST} or ProTox~\cite{Drwal2014}\cite{Banerjee2018}\cite{ProTox}.
